---
title: "TSUNAMIS - INTERNATIONAL"
author: "Emmanuel Rodríguez Silva"
date: "6/10/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. TSUNAMIS AROUND THE WORLD - DATA PROVIDED BY THE NOAA (US)


This document contains some practice using datasets that were acquired in SQL (Structured Query Language).

> This project have the following objectives:
>
>> 1. Practice R programming
>> 2. Combine the usage of SQL and R
>> 3. Make better graphics
>

```{r loading_libraries, message=FALSE, warning=FALSE}
library(palmerpenguins)
library(ggplot2)
library(tidyverse)
library(dplyr)
library(lubridate)
library(here)     # referencing files easier
library(skimr)    # simplify data cleaning task (summarizing)
library(janitor)

```


```{r reading_data}

tsunamis <- read.csv(file = "tsunamis.csv", header = T)
head(tsunamis)

str(tsunamis)

```


```{r exploring_data}
str(tsunamis)
colnames(tsunamis)

```




Now, let's organize our dataset in order to create beautiful dataviz

```{r organize}
 tsunamis %>% group_by(country) %>% arrange(country)
 
 tsunamis_impact <- tsunamis %>% group_by(country) %>%
  summarize(mean_deaths = round(mean(deaths)), max_deaths = max(deaths), min_deaths = min(deaths)) %>%
  arrange(desc(mean_deaths)) %>% head(n = 10)

```


Now that we have ordered and cleaned our data, let's do some visualizations

```{r data_viz}
mi_col <- c("deepskyblue3")

ggplot(data = tsunamis_impact, aes(x = mean_deaths, y = reorder(country, mean_deaths))) +
  # I ll use colums to graph
  geom_col(fill = mi_col, col = "black")+
  # adding maximum values in order to show discrepancies
  geom_text(aes(label = max_deaths), hjust = -1, fontface = "italic", size = 3)+
  # ADDING LABELS
  labs(title = "Deaths caused by tsunamis per country",
      subtitle = "Most deadly tsunamis since 2000 B.C.",
      caption = "Source: The Global Historical Tsunami Database (NOAA, 2020)",
      y = "Country", x = " Deaths")+
  scale_x_continuous(limits = c(0, 40000.5), breaks = seq(0, 40000, 10000)) +
  theme_gray()+
  # CUSTOMIZING LABELS
  theme(plot.title = element_text(family = "Times", face = "bold", size = 19, hjust = 0.5, vjust = 2.5),
        plot.subtitle = element_text(family = "Times", size = 15, hjust = 0.5, vjust = 1.5),
        plot.caption = element_text(size = 13, vjust = -1),
        axis.title = element_text(family = "Times", face = "bold", size = 16))


```


As we can infer from the plot above is that mean does not represents well the deadly
effect of tsunamis, because some countries had just one tsunami, and it caused many deaths




## 2. Student Mental Health Project (Kaggle)


This time I will explore data downloaded from Kaggle, I'm interested in investigating the
mental health of university students

> My objectives are:
>
>> 1. Data cleaning and organize it
>> 2. Exploring difference by sex, major, age
>> 3. Elaborate some data viz
>

```{r loading_data_health}
mental <- read.csv(file = "student_mental_health.csv", header = T)
head(mental)

```



I want to explore the differences between sex


```{r organize_data}
colnames(mental)

mental_2 <- mental %>% mutate(sex = Choose.your.gender, anxiety = Do.you.have.Anxiety., depression = Do.you.have.Depression., course = What.is.your.course.) %>% select(sex, anxiety, Age, depression) %>%
  group_by(sex) %>% summarize(how_many_have_anxiety = sum(anxiety == "Yes"), how_many_depressed = sum(depression == "Yes"))

colnames(mental_2)
  
#  group_by(sex)

```


```{r graphics}
my_colors <- c("deeppink3", "deepskyblue4")

ggplot(data = mental_2, aes(x = sex, y = how_many_have_anxiety)) +
  geom_col(fill = my_colors, col = "black", width = 0.4)+
  scale_y_continuous(limits = c(0, 40.5), breaks = seq(0, 40, 10))+
  labs(title = "Anxiety in University Students",
       subtitle = "Differences in mental health between sex",
       caption = "Data set was collected by a survey conducted by Google forms",
       x = "Sex", y = "Number of people")+
  theme(plot.title = element_text(family = "Times", face = "bold", size = 19, hjust = 0.5, vjust = 2.5),
        plot.subtitle = element_text(family = "Times", size = 15, hjust = 0.5, vjust = 1.5),
        plot.caption = element_text(family = "Times", size = 11),
        axis.title = element_text(family = "Times", size = 13, face = "bold"))+
  theme(aspect.ratio = 0.8)


```




```{r lets_explore_courses}

mental_3 <- mental %>% mutate(sex = Choose.your.gender, anxiety = Do.you.have.Anxiety., depression = Do.you.have.Depression., course = What.is.your.course.) %>% select(sex, anxiety, Age, depression, course)

mental_4 <- mental_3 %>% distinct(course)
# there is nothing interesting about courses, we need more data

```





##  Let's construct a linear regression model based on information regarding experience and salary


> I have just 2 main porpuses:
>
>> to recall the process of modelling Linear Regression models and,
>> to interpret the results
>


```{r loading_info}

salary <- read.csv(file = "Salary_dataset.csv", header = T)
head(salary, n = 8)

```


Let's observe a scatterplot to observe if there is an association between these two variables

```{r scatter_plot}
ggplot(data = salary, aes(x = YearsExperience, y = Salary))+
  geom_point(cex = 2, col = "mediumorchid2")+
  labs(title = "Salary vs. Experience",
       caption = "Data provided by Allena Venkata from her Kaggle  account (2023)",
       x = "Experience in Years")+
  theme_grey()+
  theme(plot.title = element_text(family = "Times", face = "bold", size = 19, hjust = 0.5, vjust = 2.5),
        plot.caption = element_text(family = "Times", size = 11),
        axis.title = element_text(family = "Times", size = 13, face = "bold"))
  


```

Now, I have seen some evidence about the relationship between salary and experience, I need some statistical proves, I'll apply test in order to discover the type of distribution, if data is normal, I ll apply parametric statistics, otherwise, we should use non-parametric statistics.


```{r correlation}
shapiro.test(salary$Salary)
shapiro.test(salary$YearsExperience)

```
According to the normality test, only the experience has a normal distribution, because P is greater than **0.05**

So, let's try to predict the salary based on experience, but first of all, I need to make sure there is
significance relationship between them


```{r correlation_coefficient}
cor.test(salary$Salary, salary$YearsExperience)


```

After applying the **pearson's correlation coefficient**, I noticed that exist a strong positive correlation between our variables, having said that, I will proceed with the elaboration of the linear model.



```{r linear_model}


model_salary <- lm(Salary ~ YearsExperience, data = salary)  
summary(model_salary)


plot(model_salary)

```

The previous model indicates that, for one (unit = year)increment in experience, the salary expected will increase **$9,450.00**  Knowing that, We can predict a salary based on how many years of experience a person has.


```{r predicting_manually}
# manually predicting salary according with coefficients model

print("y =24848.2 + 9450.0 * 10")


# calculating salary for a person with 10 years of experience

salary_predicted = 24848.2 + 9450.0 * 10

salary_predicted
```


```{r predinting_with_fucntion}
value <- data.frame(YearsExperience = 10)


predict(model_salary, value)

```



```{r ggplot_lm}
ggplot(data = salary, aes(x = YearsExperience, y = Salary)) +
  geom_point() +
  geom_smooth(method='lm')+
  geom_text(x = 8.5, y = 48000, label = "y = 24848.2 + 9450.0 * experience + e \n  y = b0 + b1*x + e")+
  labs(title = "Linear Regression (Predicting Salary)",
       caption = "Source: Salary Dataset on Kaggle by Allena Venkata (2023)")+
  theme_grey()+
  theme(plot.title = element_text(family = "Times", face = "bold", size = 19, hjust = 0.5, vjust = 2.5),
        plot.caption = element_text(family = "Times", size = 11),
        axis.title = element_text(family = "Times", face = "bold", size = 15))



```















